---
title: "Model Evaluation and Validation"
author: "Dr. Mine Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
    anchor_sections: FALSE
---

class: title-slide


<br>
<br>
.pull-right[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]

---

```{r echo = FALSE, message = FALSE}
library(tidyverse)
library(openintro)
library(broom)
library(modelr)
theme_set(theme_gray(base_size = 18))
```



```{r}
glimpse(babies)
```

---

```{r}
model_g <- lm(bwt ~ gestation, data = babies)
tidy(model_g)
```

---

```{r}
glance(model_g)
```

---

```{r}
glance(model_g)$r.squared
```

16% of the variation in birth weight is explained by gestation. Higher values of R squared is preferred.

```{r echo = FALSE, fig.align='center', message=FALSE, warning = FALSE, fig.height = 4}
babies %>% 
  ggplot() +
  aes(x = gestation, y = bwt) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) 
    
```

---

```{r echo = FALSE, fig.align='center', message=FALSE, warning = FALSE, fig.height = 4}
babies %>% 
  ggplot() +
  aes(x = gestation, y = bwt) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) 
    
```

---

```{r}
model_gsa <- lm(bwt ~ gestation + smoke + age, data = babies)

```


---

## Adjusted R-Squared

.pull-left[
```{r}
glance(model_g)

```
]

.pull-right[
```{r}
glance(model_gsa)
```
]


---

```{r}
babies %>% 
  add_predictions(model_g) %>% 
  add_residuals(model_g) %>% 
  select(bwt, pred, resid)
```


---

```{r}
babies %>% 
  add_predictions(model_gsa) %>% 
  add_residuals(model_gsa) %>% 
  select(bwt, pred, resid)
```

---

## Root Mean Square Error

$RMSE = \sqrt{\frac{\Sigma_{i=1}^n(y_i-\hat y_i)^2}{n}}$
---

```{r}
modelr::rmse(model_gsa, babies)
```

```{r}
modelr::rmse(model_g, babies)
```

---

```{r}
model_full <- lm(bwt ~ gestation + parity + age +
                   height + weight + smoke, 
                 data = babies)
```

```{r}
modelr::rmse(model_full, babies)
```

```{r}
glance(model_full)$r.squared
```

Can we keep adding all the variables and try to get an EXCELLENT model fit?

---

## Overfitting

- We are fitting the model to sample data.

- Our goal is to understand the population data.

- If we make our model too perfect for our sample data, the model may not perform as well with other sample data from the population.

- In this case we would be overfitting the data.

- We can use **model validation** techniques.

---

## Splitting the Data (Train vs. Test)

```{r message=FALSE}
set.seed(12345)
babies_split <- rsample::initial_split(babies) ## 75% to 25% split
```

--

```{r}
babies_train <- rsample::training(babies_split)
dim(babies_train)
```

--

```{r}
babies_test <- rsample::testing(babies_split)
dim(babies_test)
```

---

```{r}
model_gsa_train <- lm(bwt ~ gestation + smoke + age, data = babies_train)

model_gsa_test <- lm(bwt ~ gestation + smoke + age, data = babies_test)
```

---

.pull-left[
```{r}
glance(model_gsa_train)
```
]

.pull-right[
```{r}
glance(model_gsa_test)
```
]


---

.pull-left[
```{r}
modelr::rmse(model_gsa_train, babies_train)
```
]

.pull-right[
```{r}
modelr::rmse(model_gsa_test, babies_test)
```
]

