---
title: "The Pipe Operator"
author: "Dr. Mine Dogucu"
output: 
  xaringan::moon_reader:
    css: ["slide-style.css"]
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
---

class: title-slide

```{r echo = FALSE}
library(magrittr)
library(fabricerin)
```


<br>
<br>
.right-panel[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]

---

class: inverse middle

.font50[Three solutions to a single problem]

---

class: middle

What is the average of 4, 8, 16 approximately?

---

class: middle

1.What is the average of **4, 8, 16** approximately?

---

class: middle

2.What is the **average** of 4, 8, 16 approximately?

---

class: middle

3.What is the average of 4, 8, 16 **approximately**?

---

class: middle inverse

.font50[Solution 1: Functions within Functions]

---

```{r}
c(4, 8, 16)
```

--

<hr>

```{r}
mean(c(4, 8, 16))
```

--

<hr>

```{r}
round(mean(c(4, 8, 16)))
```

---

class: middle

**Problem with writing functions within functions**

Things will get messy and more difficult to read and debug as we deal with more complex operations on data.

---

class: middle inverse

.font50[Solution 2: Creating Objects]

---

class: middle

```{r}
numbers <- c(4, 8, 16)
```

--

<hr>

```{r}
avg_number <- mean(numbers)
```

--

<hr>

```{r}
round(avg_number)
```

---

class: middle

**Problem with creating many objects**

We will end up with too many objects in `Environment`. 

---

class: middle inverse

.font50[Solution 3: The (forward) Pipe Operator %>% ]

---

class: middle

.font75[Shortcut: <br>Ctrl (Command) + Shift + M]

---
class: middle

.pull-left[
```{r}
c(4, 8, 16) %>% 
  mean() %>% 
  round()
```
]

.pull-right[

Combine 4, 8, and 16 `and then`  
Take the mean   `and then`  
Round the output

]


---

.pull-left[

```{r eval = FALSE}
c(4, 8, 16) %>% 
  mean() %>% 
  round()
```
]

.pull-right[

```{r echo = FALSE}
fabric_drawing(cid = "canvas2", 
               cwidth = 520, 
               cheight = 550, 
               cfill = "#e2d7c4", 
               drawingWidth = 2, 
               gumSize = 32)


```
]

---

class: middle

The output of the first function is the first argument of the second function.

---

Do you recall composite functions such as $f \circ g(x)$?

--

Now we have $f \circ g \circ h (x)$ or `round(mean(c(4, 8, 16)))`

--

.pull-left[

```{r eval = FALSE}
h(x) %>% 
  g() %>% 
  f()
```

]

.pull-right[

```{r eval = FALSE}
c(4, 8, 16) %>% 
  mean() %>% 
  round()
```

]