---
title: "Indicator Variables"
author: "Dr. Mine Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
    anchor_sections: FALSE
---

class: title-slide


<br>
<br>
.pull-right[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]

---


```{r echo = FALSE, message = FALSE}
library(tidyverse)
library(openintro)
library(broom)
theme_set(theme_gray(base_size = 18))
```

#### Data `babies` in `openintro` package

```{r echo = FALSE}
glimpse(babies)
```


---

class: middle

<div align = "center">

| y | Response    | Birth weight | Numeric |
|---|-------------|-----------------|---------|
| x | Explanatory | Smoke           | Categorical |


---

## Notation

$y_i = \beta_0 +\beta_1x_i + \epsilon_i$

$\beta_0$ is y-intercept  
$\beta_1$ is slope  
$\epsilon_i$ is error/residual  
$i = 1, 2, ...n$ identifier for each point


---

```{r}
model_s <- lm(bwt ~ smoke, data = babies)
tidy(model_s)
```



--

$\hat {y}_i = b_0 + b_1 x_i$

$\hat {\text{bwt}_i} = b_0 + b_1 \text{ smoke}_i$

$\hat {\text{bwt}_i} = 123 + (-8.94\text{ smoke}_i)$

---

class: middle

.pull-left[

#### Expected bwt for a baby with a non-smoker mother

$\hat {\text{bwt}_i} = 123 + (-8.94\text{ smoke}_i)$

$\hat {\text{bwt}_i} = 123 + (-8.94\times 0)$

$\hat {\text{bwt}_i} = 123$

$E[bwt_i | smoke_i = 0] = b_0$


]

--

.pull-right[
#### Expected bwt for a baby with a smoker mother

$\hat {\text{bwt}_i} = 123 + (-8.94\text{ smoke}_i)$

$\hat {\text{bwt}_i} = 123 + (-8.94\times 1)$

$\hat {\text{bwt}_i} = 114.06$

$E[bwt_i | smoke_i = 1] = b_0 + b_1$

]

---

```{r}
confint(model_s)
```

Note that the confidence interval for the "slope" does not contain 0 and all the values in the interval are negative. 

