---
title: "Multiple Linear Regression"
author: "Dr. Mine Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
    anchor_sections: FALSE
---

class: title-slide


<br>
<br>
.pull-right[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]

---


```{r echo = FALSE, message = FALSE}
library(tidyverse)
library(openintro)
library(broom)
theme_set(theme_gray(base_size = 18))
```

#### Data `babies` in `openintro` package

```{r echo = FALSE}
glimpse(babies)
```

---


class: middle

<div align = "center">

| y | Response    | Birth weight | Numeric |
|---|-------------|-----------------|---------|
| $x_1$ | Explanatory | Gestation           | Numeric |
| $x_2$ | Explanatory | Smoke           | Categorical |

---

## Notation

$y_i = \beta_0 +\beta_1x_{1i}  + \beta_2x_{2i} + \epsilon_i$

$\beta_0$ is intercept  
$\beta_1$ is the slope for gestation   
$\beta_2$ is the slope for smoke 
$\epsilon_i$ is error/residual  
$i = 1, 2, ...n$ identifier for each point


---

```{r}
model_gs <- lm(bwt ~ gestation + smoke, data = babies)
tidy(model_gs)
```

--

Expected birth weight for a baby who had 280 days of gestation with a smoker mother

$\hat {\text{bwt}_i} = b_0 + b_1 \text{ gestation}_i + b_2 \text{ smoke}_i$

$\hat {\text{bwt}_i} = -0.932 + (0.443 \times 280) + (-8.09 \times 1)$

---

```{r}
babies %>% 
  modelr::add_predictions(model_gs) %>% 
  select(bwt, gestation, smoke, pred)
```

---

```{r}
babies %>% 
  modelr::add_predictions(model_gs) %>% 
  modelr::add_residuals(model_gs) %>% 
  select(bwt, gestation, smoke, pred, resid)
```


