---
title: "Data Types"
author: "Dr. Mine Dogucu"
output: 
  xaringan::moon_reader:
    css: slide-style.css
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
---

class: title-slide

<br>
<br>
.right-panel[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]

---

```{r echo = FALSE, message=FALSE}
library(fivethirtyeight)
library(openintro)
library(tidyverse)
```


```{r}
glimpse(candy_rankings)
```

---

```{r}
glimpse(mariokart)
```

---

class: middle



`character`: takes string values (e.g. a person's name, address)  
--

`integer`: integer (single precision)  
--

`double`: floating decimal (double precision)   
--

`numeric`: integer or double  
--

`factor`: categorical variables with different levels  
--

`logical`: TRUE (1), FALSE (0)  

---
class: inverse middle

As a data scientist it is .font30[**your**] job to check the type(s) of data that you are working with. Do .font30[**not**]  assume you will work with clean data frames, with clean names, labels, and types.

---


```{r echo = FALSE}
library(titanic)
```

```{r}
glimpse(titanic_train)
```


