---
title: "First Steps in R"
author: "Dr. Mine Dogucu"
output: 
  xaringan::moon_reader:
    lib_dir: libs
    css: "slide-style.css"
    seal: false
    nature:
      ratio: 16:9
      highlightStyle: "pygments"
      highlightLines: true
      highlightLanguage: "r"
---

class: title-slide

```{r child = "../setup.Rmd"}
```

<br>
<br>
.right-panel[ 

# `r rmarkdown::metadata$title`
## `r rmarkdown::metadata$author`
]

---

class: middle


We will use RStudio on ICS Hub.

--

You can access Hub at [hub.ics.uci.edu](https://hub.ics.uci.edu).

--

You will need ICS username and password. If you do not have one you can follow the instructions on the [help page](https://www.ics.uci.edu/~lab/students/acct_activate.php). 

---

class: middle center


#### Starting RStudio

<video width="80%" height="45%%" align = "center" controls>
  <source src="screencast/1-open-rstudio.mp4" type="video/mp4">
</video>



---

class: middle center inverse

.font150[hello woRld]



---

class: middle center

<video width="80%" height="45%%" align = "center" controls>
  <source src="screencast/1-hello-world.mp4" type="video/mp4">
</video>



---

class: inverse middle center

.font100[R review]

---

class: middle 

## Object assignment operator

```{r}
birth_year <- 1950
```


--

|                            | Windows        | Mac              |
|----------------------------|----------------|------------------|
| Shortcut     | Alt + -        | Option + -       |


---

class: 

## R is case-sensitive


```{r error=TRUE}

my_age <- 2020 - birth_year

My_age
```

--


---


class: middle

If something comes in quotes, it is not defined in R. 


```{r}

ages <- c(25, my_age, 32)

names <- c("Menglin", "Mine", "Rafael")

data.frame(age = ages, name = names)
```

---


## Vocabulary

```{r eval=FALSE}
do(something)
```

`do()` is a function;   
`something` is the argument of the function.
--

```{r eval=FALSE}
do(something, colorful)
```

`do()` is a function;   
`something` is the first argument of the function;   
`colorful` is the second argument of the function.

---

class: middle

## Getting Help

In order to get any help we can use `?` followed by function (or object) name. 

```{r eval=FALSE}
?c
```


---

## tidyverse_style_guide

>canyoureadthissentence?


--
    
.pull-right[

```{r eval = FALSE}
age <- c(6, 9, 15)

data.frame(age_kid = age)
```

]

--

.pull-left[

After function names do not leave any spaces.

Before and after operators (e.g. <-, =) leave spaces. 

Put a space after a comma, **not** before. 

Object names are all lower case, with words separated by an underscore.

]



---

class: middle center


#### RStudio Setup

<video width="80%" height="45%%" align = "center" controls>
  <source src="screencast/1-rstudio-setup.mp4" type="video/mp4">
</video>